How does an Experimental Software Engineering Lab Work? An Ethnographic Study

### 2023-07-23 tex/abstract.tex
some more changes
A couple of more changes.

\begin{abstract}
\textbf{Context:} Replication is complex in Experimental Software Engineering (ESE). Replication packages and families of experiments have been applied with limited success. We wonder whether the problems are due to formal issues, e.g., under-specification, miscommunication, etc., or intrinsic reasons, i.e., the researchers' needs are not fulfilled by the existing replication procedures.
\textbf{Objective:} Find out how ESE researchers conduct experiments in practice.
\textbf{Method:} Ethnographical study with an experimental research group.
\textbf{Results:} We have created conceptual and process models representing how experimentation, replication, and synthesis are conducted in the target research group. These models fit the mainstream procedures and terminology at a high level but depart at lower, specialized levels. The experimental process carried out in the group differs from folk knowledge and textbooks in (1) the number and diversity of activities, (2) the existence of different roles, (3) the granularity of the concepts, and (4) the viewpoints that different sub-areas or families of experiments have about the experimental process.
\textbf{Conclusions:} The differences between the actual lab processes vs. the idealized versions probably harm knowledge transfer and make replication harder. We will conduct further research with more research groups, e.g., using a survey, to explore whether our findings have general applicability.
\end{abstract}

>>>>

\begin{abstract}
\textbf{Context:} Replication is complex in Experimental Software Engineering (ESE). Replication packages and families of experiments have been applied with limited success. We wonder whether the problems are due to formal issues, e.g., under-specification, miscommunication, or intrinsic reasons, i.e., the existing replication procedures do not fulfill the researchers' needs.
\textbf{Objective:} Find out how ESE researchers conduct experiments in practice.
\textbf{Method:} Ethnographical study with an experimental research group.
\textbf{Results:} We have created conceptual and process models representing experimentation, replication, and synthesis in the target research group. These models fit the mainstream procedures and terminology at a high level but depart at lower, specialized levels. The experimental process carried out in the group differs from folk knowledge and textbooks in (1) the number and diversity of activities, (2) the existence of different roles, (3) the granularity of the concepts, and (4) the viewpoints that different sub-areas or families of experiments have about the experimental process.
\textbf{Conclusions:} The differences between the actual lab processes vs. the idealized versions probably harm knowledge transfer and make replication harder. We will conduct further research with more research groups, e.g., using a survey to explore whether our findings have general applicability.
\end{abstract}

### 2023-07-23 tex/introduction.tex
some more changes
A couple of more changes.

This paper aims to find out \textbf{how experimental researchers conduct experiments in practice}. We applied an ethnographic method \cite{Sharp-2016-Ethnographic-Studies-ESE} within an experienced experimental research group. We observed the researchers in their lab for two years. We created conceptual and process models to represent the concepts they use and the activities that they perform on a day-to-day basis. These models fit the community's procedures and terminology at a high level. Still, the models show particularities at lower levels: (1) The number and diversity of activities, (2) the existence of different roles, (3) the granularity of the concepts used by the roles, and (4) the viewpoints that different sub-areas or families of experiments have about the overall process.

@@

\begin{itemize}
  \item Identifying activities that depart from, or are not described in, mainstream experimentation procedures. 
  \item Characterizing roles, i.e., groups of experimental activities performed by the same person, in ESE research.
  \item Evidence of the existence of tacit knowledge in ESE, which, in turn, leads to discrepancies between experimentation protocols and actual practice.
  \item Conceptual and process models that represent how experimentation is conducted \textit{in practice} in an experimental research group.
\end{itemize}

Finally, we wish to open a discussion in the ESE community: 
\begin{itemize}
  \item Should those differential characteristics in experimentation in software engineering be disregarded as inherently contextual?
  \item  \textbf{Alternatively}: Are they more common than it may seem, so they should be further investigated and eventually integrated into mainstream practice?
\end{itemize}

>>>>

This paper aims to find out \textbf{how experimental researchers conduct experiments in practice}. We applied an ethnographic method \cite{Sharp-2016-Ethnographic-Studies-ESE} within an experienced experimental research group. We observed the researchers in their lab for two years. We created conceptual and process models to represent the concepts they use and the activities they perform daily. These models fit the community's procedures and terminology at a high level. Still, the models show particularities at lower levels: (1) The number and diversity of activities, (2) the existence of different roles, (3) the granularity of the concepts used by the roles, and (4) the viewpoints that different sub-areas or families of experiments have about the overall process.

@@

\begin{itemize}
  \item Identifying activities that depart from or are not described in mainstream experimentation procedures. 
  \item Characterizing roles, i.e., groups of experimental activities performed by the same person, in ESE research.
  \item Evidence of the existence of tacit knowledge in ESE, which, in turn, leads to discrepancies between experimentation protocols and actual practice.
  \item Conceptual and process models that represent how experimentation is conducted \textit{in practice} in an experimental research group.
\end{itemize}

Finally, we wish to open a discussion in the ESE community: 
\begin{itemize}
  \item Should those differential characteristics in experimentation in software engineering be disregarded as inherently contextual?
  \item  \textbf{Alternatively}: Are they more common than they may seem, so they should be further investigated and eventually integrated into mainstream practice?
\end{itemize}
